---
output: github_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/images/"
)
```

# Fit seeded LDA using quanteda

A function to fit seeded LDA model for semi-supervised topic classification using **topicmodels** and **quanteda**. The code is from [*Making a topic dictionary for semi-supervised classification of the UN speeches*](https://koheiw.net/wp-content/uploads/2019/06/Speech-classification-06-QTA-Dub.pdf).

## Install

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("koheiw/quanteda.seededlda")
```

## Example

The corpus and seed words in this example are from [*Conspiracist propaganda: How Russia promotes anti-establishment sentimentonline?*](https://koheiw.net/wp-content/uploads/2019/06/Sputnik-05-ECPR.pdf)

```{r message=FALSE}
require(quanteda)
require(topicmodels)
require(quanteda.seededlda)
```

```{r}
corp <- readRDS("inst/extdata/data_corpus_sputnik.RDS")
toks <- tokens(corp, remove_punct = TRUE)
dfmt <- dfm(toks) %>% 
    dfm_select("^[A-Za-z]+$", valuetype = "regex") %>% 
    dfm_remove(stopwords('en')) %>% 
    dfm_trim(min_termfreq = 0.90, termfreq_type = "quantile", 
             max_docfreq = 0.1, docfreq_type = "prop")
```

```{r}
dict <- dictionary(file = "inst/extdata/topics.yml")
print(dict)
```

```{r}
slda <- textmodel_seededlda(dfmt, dict, residual = TRUE)
term <- terms(slda, 20)
colnames(term) <- c(names(dict), "else")
print(term)
```

```{r}
topic <- table(topics(slda))
names(topic) <- c(names(dict), "else")
print(topic)
```

