---
output: github_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/images/"
)
```

# seededlda: the package for semi-supervised topic modeling

<!-- badges: start -->

[![CRAN
Version](https://www.r-pkg.org/badges/version/seededlda)](https://CRAN.R-project.org/package=seededlda)
[![Downloads](https://cranlogs.r-pkg.org/badges/seededlda)](https://CRAN.R-project.org/package=seededlda)
[![Total
Downloads](https://cranlogs.r-pkg.org/badges/grand-total/seededlda?color=orange)](https://CRAN.R-project.org/package=seededlda)
[![R build
status](https://github.com/koheiw/seededlda/workflows/R-CMD-check/badge.svg)](https://github.com/koheiw/seededlda/actions)
[![codecov](https://codecov.io/gh/koheiw/seededlda/branch/master/graph/badge.svg)](https://codecov.io/gh/koheiw/seededlda)
<!-- badges: end -->

**seededlda** is an R package that implements the seeded-LDA for semi-supervised topic modeling using **quanteda**. The seeded-LDA model was proposed by [Lu et al. (2010)](https://dl.acm.org/citation.cfm?id=2119585). Until version 0.3, that packages has been a simple wrapper around the **topicmodels** package, but the LDA estimator is newly implemented in C++ using the [GibbsLDA++](http://gibbslda.sourceforge.net/) library to be submitted to CRAN in August 2020. The author believes this package implements the seeded-LDA model more closely to the original proposal.

Please see [*Theory-Driven Analysis of Large Corpora: Semisupervised Topic Classification of the UN Speeches*](https://journals.sagepub.com/doi/full/10.1177/0894439320907027) for the overview of semi-supervised topic classification techniques and their advantages in social science research.

[**keyATM**](https://github.com/keyATM/keyATM) is the latest addition to the semi-supervised topic models. The users of seeded-LDA are also encouraged to use that package.

## Installation

From CRAN:

```{r, eval=FALSE}
install.packages("seededlda")
```

From Github:

```{r, eval=FALSE}
devtools::install_github("koheiw/seededlda")
```

## Examples

Please visit the package website for examples:

- [Introduction](https://koheiw.github.io/seededlda/articles/pkgdown/basic.html): basic functions of the package
- [Distributed LDA](https://koheiw.github.io/seededlda/articles/pkgdown/distributed.html): topic modeling with parallel computing
- [Seeded LDA](https://koheiw.github.io/seededlda/articles/pkgdown/seeded.html): semi-supervised topic modeling
- [Sequential LDA](https://koheiw.github.io/seededlda/articles/pkgdown/sequential.html): sentence-level topic modeling

Please read the following papers on the algorithms.

- Watanabe, K., & Baturo, A. (2023). Seeded Sequential LDA: A Semi-Supervised Algorithm for Topic-Specific Analysis of Sentences. Social Science Computer Review. https://doi.org/10.1177/08944393231178605
- Watanabe, K. (2023). Speed Up Topic Modeling: Distributed Computing and Convergence Detection for LDA, [working paper](https://blog.koheiw.net/wp-content/uploads/2023/05/Distributed-LDA-02.pdf).


## Other publications

Please read the following papers for how to apply seeded-LDA in social science research:

Curini, Luigi and Vignoli, Valerio. 2021. [Committed Moderates and Uncommitted Extremists: Ideological Leaning and Parties’ Narratives on Military Interventions in Italy](https://doi.org/10.1093/fpa/orab016), *Foreign Policy Analysis*.

