---
title: "Introduction to seededlda"
subtitle: "The package for semi-supervised topic modeling"
output: 
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = FALSE, comment = "#>", 
                      fig.width = 8, fig.height = 4, dpi = 150, out.width = 760)
```

**seededlda** was crated mainly for semi-supervised topic modeling but it can perform unsupervised topic modeling too. I explain the basic functions of the package taking unsupervised topic modeling as an example in this page and discuss semi-supervised topic modeling in [a separate page](seeded.html).

### Preperation

We use [the corpus of Sputnik articles](https://www.dropbox.com/s/abme18nlrwxgmz8/data_corpus_sputnik2022.rds?dl=1) about Ukraine in the examples. In the preprocessing, we remove grammatical words `stopwords("en")`, email addresses `"*@*` and words that occur in more than 10% of documents `max_docfreq = 0.1` from the document-feature matrix (DFM).

```{r include=FALSE}
if (!file.exists("data_corpus_sputnik2022.rds")) {
    download.file("https://www.dropbox.com/s/abme18nlrwxgmz8/data_corpus_sputnik2022.rds?dl=1",
                  "data_corpus_sputnik2022.rds", mode = "wb")
}
```

```{r message=FALSE}
library(seededlda)
library(quanteda)

corp <- readRDS("data_corpus_sputnik2022.rds")
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE, 
               remove_numbers = TRUE, remove_url = TRUE)
dfmt <- dfm(toks) |> 
    dfm_remove(stopwords("en")) |>
    dfm_remove("*@*") |>
    dfm_trim(max_docfreq = 0.1, docfreq_type = "prop")
print(dfmt)
```

### Standard LDA

You can analyze the DFM only by setting the number of topics `k = 10` to identify. When `verbose = TRUE`, it shows the progress of the inference through iterations. It takes long to fit LDA on a large corpus, but [distributed algorithm](distributed.html) will speed up your analysis dramatically.

```{r, cache=TRUE}
lda <- textmodel_lda(dfmt, k = 10, verbose = TRUE)
```

### Topic terms

Once model is fit, you can can interpret the topics by reading the most salient words in the topics. `terms()` shows words that are most frequent in each topic at the top of the matrix.

```{r, results='asis'}
knitr::kable(terms(lda))
```

### Document topics

You can also predict the topics of documents using `topics()`. I recommend extracting the document variables from the DFM in the fitted object `lda$data` and saving the topics in the data.frame. 

```{r}
dat <- docvars(lda$data)
dat$topic <- topics(lda)
```

```{r, results='asis'}
knitr::kable(head(dat[,c("date", "topic", "head")], 10))
```

### References

- Heinrich, G. (2008). Parameter estimation for text analysis. http://www.arbylon.net/publications/text-est.pdf
- Watanabe, K., & Baturo, A. (2023). Seeded Sequential LDA: A Semi-Supervised Algorithm for Topic-Specific Analysis of Sentences. Social Science Computer Review. https://doi.org/10.1177/08944393231178605
