---
title: "Distributed LDA"
subtitle: "LDA with parallel computing"
output: 
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", 
                      fig.width = 8, fig.height = 4, dpi = 150, out.width = 760)
```

In this example, we will analyze how much the proposition of [the European Gas Demand Reduction Plan](https://ec.europa.eu/commission/presscorner/detail/en/ip_22_4608) on 20 July affected Sputnik's coverage of energy issues in the United States and the European Union.

```{r message=FALSE}
library(seededlda)
library(quanteda)
```

### Preperation

```{r}
corp <- readRDS("data_corpus_sputnik2022.rds")
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE, 
               remove_numbers = TRUE, remove_url = TRUE)
dfmt <- dfm(toks) |> 
    dfm_remove(stopwords("en")) |>
    dfm_remove("*@*") |>
    dfm_trim(max_docfreq = 1000)
print(dfmt)
```


### Distributed LDA


```{r}
lda_dist <- textmodel_lda(dfmt, k = 10, batch_size = 0.1, verbose = TRUE)
```

```{r, results='asis'}
knitr::kable(terms(lda_dist))
```

### Distributed LDA with convergence detection

```{r}
lda_auto <- textmodel_lda(dfmt, k = 10, batch_size = 0.1, auto_iter = TRUE,
                      verbose = TRUE)
```

```{r, results='asis'}
knitr::kable(terms(lda_auto))
```

### References

- Newman, D., Asuncion, A., Smyth, P., & Welling, M. (2009). Distributed Algorithms for Topic Models. The Journal of Machine Learning Research, 10, 1801–1828.
- Watanabe, K. (2023). Speed Up Topic Modeling: Distributed Computing and Convergence Detection for LDA, working paper.
