---
title: "Seeded LDA"
subtitle: "LDA for semi-supervised topic classification"
output: 
  html_document:
    toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = FALSE, comment = "#>", 
                      fig.width = 8, fig.height = 4, dpi = 150, out.width = 760)
```

In this example, we will analyze how much the proposition of [the European Gas Demand Reduction Plan](https://ec.europa.eu/commission/presscorner/detail/en/ip_22_4608) on 20 July affected Sputnik's coverage of energy issues in the United States and the European Union.

```{r message=FALSE}
library(seededlda)
library(quanteda)
library(ggplot2)
```

### Preperation

```{r}
corp <- readRDS("data_corpus_sputnik2022.rds")
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE, 
               remove_numbers = TRUE, remove_url = TRUE)
dfmt <- dfm(toks) |> 
    dfm_group() |>
    dfm_remove("*@*") |>
    dfm_remove(stopwords("en")) |>
    dfm_trim(max_docfreq = 1000)
print(dfmt)
```

We will use a [dictionary](https://github.com/koheiw/seededlda/tree/master/vignettes) of keywords in this example. 

```{r}
dict <- dictionary(file = "dictionary.yml")
print(dict)
```

### Seeded LDA

```{r}
lda_seed <- textmodel_seededlda(dfmt, dict, batch_size = 0.1, auto_iter = TRUE,
                                verbose = TRUE)
```

```{r, results='asis'}
knitr::kable(terms(lda_seed))
```

### Seeded LDA with residual topics

```{r}
lda_res <- textmodel_seededlda(dfmt, dict, residual = 2, batch_size = 0.1, auto_iter = TRUE,
                                verbose = TRUE)
```

```{r, results='asis'}
knitr::kable(terms(lda_res))
```

### References

- Lu, B., Ott, M., Cardie, C., & Tsou, B. K. (2011). Multi-aspect sentiment analysis with topic models. 2011 IEEE 11th International Conference on Data Mining Workshops, 81–88.
- Watanabe, K., & Baturo, A. (2023). Seeded Sequential LDA: A Semi-Supervised Algorithm for Topic-Specific Analysis of Sentences. Social Science Computer Review. https://doi.org/10.1177/08944393231178605


