---
title: "Sequential LDA"
subtitle: "LDA for sentence-level topic classification"
output: 
  html_document:
    toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", 
                      fig.width = 8, fig.height = 4, dpi = 150, out.width = 760)
```


```{r message=FALSE}
library(seededlda)
library(quanteda)
library(ggmatplot)
```

### Preperation

We will analyze the same [corpus](https://www.dropbox.com/s/5tlux53aukf3uqs/data_corpus_sputnik2022.rds?dl=1) as the [introduction](./introduction.html), so too the pre-processing.

```{r}
corp <- readRDS("data_corpus_sputnik2022.rds") |>
    corpus_reshape()
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE, 
               remove_numbers = TRUE, remove_url = TRUE)
dfmt <- dfm(toks) |> 
    dfm_remove("*@*") |>
    dfm_remove(stopwords("en"))
print(dfmt)
```

### Sequential LDA

```{r}
lda_seq <- textmodel_lda(dfmt, k = 10, gamma = 0.5, 
                         batch_size = 0.1, auto_iter = TRUE,
                         verbose = TRUE)
```

```{r, results='asis'}
knitr::kable(terms(lda_seq))
```

```{r}
ggmatplot(lda_seq$theta[paste0("s1104078647.", 1:100),], 
          plot_type = "line", linetype = 1,
          xlab = "Sentence", ylab = "Probability")
```


### Seeded Sequential LDA

```{r}
dict <- dictionary(file = "dictionary.yml")
print(dict)
```

```{r}
lda_seed <- textmodel_seededlda(dfmt, dict, residual = 2, gamma = 0.5, 
                                batch_size = 0.1, auto_iter = TRUE,
                                verbose = TRUE)
```

```{r, results='asis'}
knitr::kable(terms(lda_seed))
```


```{r}
ggmatplot(lda_seed$theta[paste0("s1104078647.", 1:100),], 
          plot_type = "line", linetype = 1, 
          xlab = "Sentence", ylab = "Probability")
```


### References

- Du, L., Buntine, W., Jin, H., & Chen, C. (2012). Sequential latent Dirichlet allocation. Knowledge and Information Systems, 31(3), 475–503. https://doi.org/10.1007/s10115-011-0425-1
- Watanabe, K., & Baturo, A. (2023). Seeded Sequential LDA: A Semi-Supervised Algorithm for Topic-Specific Analysis of Sentences. Social Science Computer Review. https://doi.org/10.1177/08944393231178605

